---
title: "Comparing Datasets"
author: "Conor Tompkins"
date: "January 30, 2017"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparing Datasets in R

To use this, you'll need to install:

R (https://cran.r-project.org/)
RStudio (https://www.rstudio.com/products/rstudio/download/)

Use the standard installation options

Open RStudio.

The top left area is the "Source" panel. This is where you enter and edit code.

The bottom left area is the "Console" panel. This is where the output from your code goes.

The top right area is the "Environment/History" panel. This is where you can see your data objects.

The bottom right area is the "Files/Plots/etc." panel. We won't be using this for now, but that's where your graphs and help queries show up.


In RStudio, enter this code in the Source panel. Select the code and hit CTRL+Enter to run the code and install the package we'll need for this tutorial
``` {r install packages, eval = FALSE}
install.packages("tidyverse")
```

Now load that package into R for this session
```{r load library}
library(tidyverse)
```

Create a folder somewhere called "\comparing datasets"

Download the roster_1516.csv and roster_1617 files to that folder

Set your working directory to that folder:

Top menu bar -> Session -> Set Working Directory -> Choose Directory

That tells R where to look for the files, and where to save any files you create in R

Confirm this using
```{r getwd}
getwd()
```
That's my working directory. Yours should be whatever your new folder is.

Read the CSVs into R
```{r read CSVs}
roster_1516 <- read.csv("roster_1516.csv") #THis read the the CSV into R and creates a dataframe called "roster_1516"
roster_1617 <- read.csv("roster_1617.csv")
```

Then run this to confirm R has them.


```{r confirm CSVs}
ls()
```
This lists all the data objects in your environment

Also check the Environment panel


The format of data in R is very important

If the class is not right, your code won't work

First, check the classes of the columns of roster_1516
```{r check roster_1516}
str(roster_1516)
```

The Player column (roster_1$Player) is a factor (fct). It needs to be a character (chr) so we can work with it
```{r as character 1516}
roster_1516$Player <- as.character(roster_1516$Player)
```
Now check if R thinks roster_1516$Player is a character class column
```{r is character}
is.character(roster_1516$Player)
```

Now, check the classes of the columns of roster_1617
```{r check roster_1617}
str(roster_1617)#roster_2$Player is a factor (fct). It needs to be a character (chr)
roster_1617$Player <- as.character(roster_1617$Player)
is.character(roster_1617$Player)
```

Which players from 1516 are in the roster for 1617?
```{r which match}
roster_1516[which(roster_1516$Player %in% roster_1617$Player), ]
```

Which players from 1516 are NOT (!) in the roster for 1617?
```{r which dont match}
roster_1516[which(!(roster_1516$Player %in% roster_1617$Player)), ]
```

But what about Matt Murray?
```{r matt murray}
roster_1516[grep("Murray", roster_1516$Player), ] #"grep" Searches the Player column for a string containing "Murray"
roster_1617[grep("Murray", roster_1617$Player), ]
```

That's because
```{r murray check}
"Matt Murray" == "Matthew Murray"
```

Let's add add a "check column to roster_1516
```{r add check column}
roster_1516 <- roster_1516 %>% #Make changes to roster_1516 and assign it back to roster_1516
  mutate(check = Player %in% roster_1617$Player) #Create a new column "check" that checks if the rows in Player column from roster_1516 match the rows in roster_1617
```

Now create a new dataframe with just the matches
```{r new dataframe}
new_dataframe <- roster_1516 %>% #Make changes to roster_1516 and assign it to a new object, new_dataframe
  filter(check == TRUE) #Filter only the rows where the check column is TRUE
```

You should have a new object "new_dataframe" in your environment
```{r list files 2}
ls()
```

View the first 10 rows of new_dataframe
```{r head}
head(new_dataframe)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
